---
title: "Hierarchical Clustering - Dendrogram"
author: "Drew Sandberg | drew.sandberg@gmail.com"
date: "4/25/2020"
output:
  html_document: default
  word_document: default
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(dendextend)
```

Dataset column names and sampling probabilities derived from copyrighted (@2017)data by DecisionPro, Inc. used in
MIS 722 - Marketing Engineering and Analytics lesson plan Dr. LinLin Chai, NDSU. 
Data set values are generated from sampling 0 - 10 using probabilities
```{r}
#set seed so dataset values, as calcuated from sampling probabilities, are generated with the same values each time.
set.seed(141)

#construct the dataset; i.e the "segmentation data"
#observations = customer/respondent feedback
#sample function says to generate 40 numbers between 0 and 10 where each of the numbers (i.e. 0, 1,2,3 etc.) 
#has the following probability of occuring in the sample.
dataset = data.frame(
    observation = seq(1,40,1), #generate customers 1 through 40
    variety = sample(0:10,40,
                     replace = TRUE, 
                     prob = c(0, 0, 0, 0, 0.1, 0.075, 0.1, 0.225, 0.175, 0.075, 0.25)
                     ),
    electronics = sample(0:10,40,
                         replace = TRUE, 
                         prob = c(0, 0.075, 0.05, 0.15, 0.175, 0.225, 0.225, 0.025, 0.075, 0, 0)
                         ),
    furniture = sample(0:10,40,
                       replace = TRUE, 
                       prob = c(0.1, 0.125, 0.25, 0.075, 0.075, 0.1, 0.125, 0.15, 0, 0, 0)
                       ),
    QoS = sample(0:10,40,
                 replace = TRUE, 
                 prob = c(0, 0.15, 0.2, 0.125, 0.25, 0.075, 0, 0.025, 0.05, 0.125, 0)
                 ),
    price = sample(0:10,40,
                   replace = TRUE, 
                   prob = c(0, 0.025, 0.225, 0.1, 0.125, 0.175, 0, 0.125, 0.075, 0.075, 0.075)
                   ),
    returns = sample(0:10,40,
                     replace = TRUE, 
                     prob = c(0, 0, 0.15, 0.225, 0.1, 0.25, 0.125, 0.075, 0.075, 0, 0)
                     )
  )
```

Next, normalize the values in the dataset, but omit the observation column.
The dendrogram is created with the hclust function

```{r}
#normalizing the data using scale function
#omitting the first column; normalizing only the variety, electornics... returns values
scaled_df = scale(dataset[2:7])

#hc = hierarchical clustering
#'dist' argument of hclust function computes a distance matrix between all values
hc = hclust(dist(scale(dataset[2:7]), method = "euclidean"), method = "average")
```

Next, print/plot the dendrogram
```{r}
plot(hc)
```

To better visualize the clusters, let R draw colored rectangles around the clusters -- based upon your desired number of clusters.
```{r}
#draw pretty colors around clusters; you pick the number of clusters using the k argument (e.g. k = 3)
#border defines the colors you want to use, or you can used 
plot(hc)
rect.hclust(hc, k = 3, border = 3:9)
```

Alternatively, you can define border color using an index of colors
```{r}
plot(hc)
rect.hclust(hc, k = 4, border = 2:6)
```
